Choose a License

### 2025-01-25 choose_license.qmd
Add full license texts

Having selected the licenses of your choice -- again, you might need multiple ones depending on the types of works your project contains --, we encourage you to read through the full license text (or at least a legal summary) to understand their effect. Then, you can record the license of existing content and apply the licenses of your own contributions. Mostly, this just means indicating which file or folder is covered by which license, usually in the project's README (among other places), whose creation will be discussed [later](make_readme.qmd). Notably, no sort of registration is required.

::: {#cau-license-versions .callout-caution}
### License Versions Are Important

You may have noticed that we mostly refer to licenses using a name _and_ a version number. This is because the organizations that created the licenses sometimes publish updated versions to accommodate for developments in copyright law and the communities that use the licenses. For example, the Creative Commons licenses (that start with `CC`) were first published in 2002. Since then, the possibility to relicense under later ([v2.0](https://creativecommons.org/weblog/entry/4216#:~:text=Share%20Alike%20Across%20Borders), 2004) and compatible licenses ([v3.0](https://creativecommons.org/weblog/entry/7249), 2007) has been added, a 30-day window to correct license violations has been established to combat [copyleft trolls](https://commons.wikimedia.org/wiki/Commons:Copyleft_trolling), and _sui generis_ database rights are covered explicitly ([v4.0](https://creativecommons.org/weblog/entry/40768), 2013). There are many more [subtle differences between license versions](https://wiki.creativecommons.org/wiki/License_Versions), therefore it is important to indicate which license version exactly one is referring to, as the license of a work does not "update" automatically. CC0, officially introduced in 2009, is still at version [1.0](http://www.creativecommons.org/weblog/entry/13304).

For the AGPLv3 it is even recommended to state whether a work is licensed under exactly the indicated version of the license or, alternatively, also under newer versions of the license [@Stallman2022Version].
:::

For example, the `apaquarto` extension that you included in your project is a work by others.^[Unless you are the author of the `apaquarto` extension.] You need to indicate its license so that others know what they are allowed to do -- and, of course, you need to comply with any terms yourself, such as retaining the copyright notice.^[In this particular case, `apaquarto` is licensed under [CC0\ 1.0](https://creativecommons.org/publicdomain/zero/1.0/), which does not require you to retain a copyright notice. In fact, it's an extremely permissive license.] In contrast, if it were not for this tutorial, the manuscript would contain your own work and you would need to indicate under which license you provide it to others.

@@

Regardless of how exactly the licenses are added to the project, this is a good opportunity to verify one last time that all third party content is provided to you under a free/open license and that you comply with it. Please add a license to your project now, either creating a file `LICENSE.txt` or following the REUSE standard.

## Wrap-up

If you would like to learn more about copyright and licenses you might find the following resources interesting:

- "Open Content -- A Practical Guide to Using Creative Commons Licences" by @Kreutzer2014
- "Creative Commons Certificate for Educators, Academic Librarians, and Open Culture" by @CC2024Certificate
- "Freie Software -- Zwischen Privat- und Gemeineigentum" by @Grassmuck2004

@@

  type --"Software"--> code_sa("Attribution?<br>Add license text?<br>State changes?<br>Disclose code?<br>Copyleft?")
  type --"Writing, image, audio, video"--> nocode_cc("Attribution?<br>State changes?<br>Anti-DRM?<br>Copyleft?")

  code_sa --"Attribution &<br>Add license text &<br>State changes"--> apache["Apache 2.0"]
  code_sa --"Attribution &<br>Disclose code &<br>Weak copyleft"--> mpl["MPL 2.0"]
  code_sa --"Attribution &<br>Add license text &<br>State changes &<br>Disclose code &<br>Strong copyleft"--> agpl["AGPLv3"]

  nocode_cc --"Neither"--> cc0["CC0 1.0"]
  nocode_cc --"Attribution &<br>State changes &<br>Anti-DRM"--> cc_by["CC BY 4.0"]
  nocode_cc --"Attribution &<br>State changes &<br>Anti-DRM &<br>Copyleft"--> cc_by_sa["CC BY-SA 4.0"]

@@

  cc0_content_metadata --> cc0_db["CC0 1.0"]
  switch_license --"CC0 or<br>non-CC license"--> sa("Attribution?<br>Anti-DRM?<br>Copyleft?")
  switch_license --"CC BY or<br>CC BY-SA"--> same["<em>Same license for DB</em>"]

  sa --"Neither"--> cc0_db
  sa --"Attribution &<br>Anti-DRM &<br>Copyleft"--> odbl["ODbL 1.0"]

>>>>

Having selected the license(s) of your choice, we encourage you to read through the full license text (or at least a legal summary) to understand the legal effect. The following applies to all licenses mentioned here:

- Applying the license is irrevocable. You may stop distributing your work, but everybody who already received it may continue to share it.
- As an author applying the license, you are not bound by the license terms. You cannot, however, give others an exclusive license to your content anymore (as some journals needlessly demand, see @wrn-exclusive-rights).
- By applying the license you share with others your right to use your work for commercial purposes, including selling it.
- If you apply the license to an image, the license [may also apply](https://commons.wikimedia.org/wiki/Commons:Same_work) to higher resolution version of your image.
- If you apply a copyleft license to an image, a document in which the image is used may not constitute a derivative work and, as such, may not be affected by the copyleft condition.
- If you are a member of a collecting society^[in Germany: _Verwertungsgesellschaft_] or if the work was created as part of your job, you may not be able to apply the license. Check with your collecting society or your employer in that case.

::: {#nte-situation-lmu .callout-note collapse="true"}
### For Researchers at LMU Munich

For Creative Commons licenses, the following is additionally important to know:

- By applying the license you waive any moral, publicity, privacy, or other similar personality rights to the extent necessary to use the work.
- The attribution requirement may be implemented "in any reasonable manner", possibly not conforming to your precise expectations.

Applying a license to a work mostly just means indicating which file or folder is covered by which license(s) (no sort of registration is required). There are multiple standards on how to do so:

- You have already learned about the file `LICENSE.txt`. Either it contains the full text of one license that applies to the whole project folder -- this is a practice propagated by GitHub, which provides instructions for and comparisons of many licenses via [ChooseALicense.com](https://choosealicense.com/licenses/). Alternatively, it details which license applies to what and the invidual licenses are stored in the same or in different files.
- A summary of the licensing situation is often given in a `README` file, whose creation will be discussed [later](make_readme.qmd).
- If only parts of a file are covered by a certain license, then it might make sense to indicate this directly in the file.
- The license of embedded elements (e.g., images) is either given directly beneath to where they are embedded, at the end of the respective page, or collectively on an separate page.
- Individual programming languages also have their own way of stating which license a package is distributed under. For R packages, this is usually set by the field `License` in the file `DESCRIPTION` [@Wickham2023PackagesLicense].
- Software projects often indicate the license in every file, following the [REUSE](https://reuse.software/) specification to make the choice of license machine-readable (explained in @nte-reuse).
- The following two metadata standards for increasing citability also allow to indicate the license:
  - [CodeMeta](https://codemeta.github.io/): All information are stored in a file called `codemeta.json`, which can be created [on their website](https://codemeta.github.io/codemeta-generator/). Alternatively, one can use the R package [`codemetar`](https://docs.ropensci.org/codemetar/) to create this file.
  - [Citation File Format](https://citation-file-format.github.io/) (CFF): Here, everything is stored in a file called `CITATION.cff`. It is used by GitHub to directly show a recommended citation for a repository. The file can be created [on their website](https://citation-file-format.github.io/cff-initializer-javascript/). Alternatively, the R package [`cffr`](https://cran.r-project.org/package=cffr) may be used for that purpose.

@@

The manuscript you downloaded for this tutorial is made available to you under [CC0\ 1.0](https://creativecommons.org/publicdomain/zero/1.0/) as is all code in this tutorial (e.g., for creating the data dictionary). Therefore, you can apply any license to your modified version. Please apply the license of your choice now, using any of the methods mentioned above (except the `README`, which we will cover later).

::: {#tip-manuscript-licensetxt .callout-tip collapse="true"}
### Addition to `LICENSE.txt` (Solution)

In the following example, the edited manuscript files have been multi-licensed under [AGPLv3](https://www.gnu.org/licenses/agpl-3.0.html) (or later) or (at option of the reuser) under [CC\ BY-SA\ 4.0](https://creativecommons.org/licenses/by-sa/4.0/), while the data dictionary and its code are made available under CC0\ 1.0.

```{.txt .code-overflow-wrap filename="LICENSE.txt"}
The Quarto documents stored in "Manuscript.qmd", "Manuscript.tex", and "Manuscript.pdf" by Josephine Zerna, Christoph Scheffel, and <YOUR NAME> are available under AGPLv3 (or later) <https://www.gnu.org/licenses/agpl-3.0.html> or (at your option) under CC BY-SA 4.0 <https://creativecommons.org/licenses/by-sa/4.0/>.

The data dictionary (stored in "data_dictionary.html") and the code used to create it (stored in "create_data_dictionary.R") by Florian Kohrt and <YOUR NAME> are available under CC0 1.0 <https://creativecommons.org/publicdomain/zero/1.0/>.
```

As the AGPLv3 requires reusers adding the [full license text](https://www.gnu.org/licenses/agpl-3.0.txt), we recommend to also copy it to the project folder (if not already in there).
:::

@@

  type --"Software"--> code_sa("Attribution?<br>Indicate license?<br>Add license text?<br>State changes?<br>Disclose code?<br>Copyleft?")
  type --"Writing, image, audio, video"--> nocode_cc("Attribution?<br>Indicate license?<br>State changes?<br>Anti-DRM?<br>Copyleft?")

  code_sa --"Attribution &<br>Add license text &<br>State changes"--> apache["Apache 2.0"]
  code_sa --"Attribution &<br>Indicate license &<br>Disclose code &<br>Weak copyleft"--> mpl["MPL 2.0"]
  code_sa --"Attribution &<br>Add license text &<br>State changes &<br>Disclose code &<br>Strong copyleft"--> agpl["AGPLv3"]

  nocode_cc --"Neither"--> cc0["CC0 1.0"]
  nocode_cc --"Attribution &<br>Indicate license &<br>State changes &<br>Anti-DRM"--> cc_by["CC BY 4.0"]
  nocode_cc --"Attribution &<br>Indicate license &<br>State changes &<br>Anti-DRM &<br>Copyleft"--> cc_by_sa["CC BY-SA 4.0"]

@@

  cc0_content_metadata --> cc0_db["CC0 1.0"]
  switch_license --"CC0 or<br>non-CC license"--> sa("Attribution?<br>Indicate license?<br>Anti-DRM?<br>Copyleft?")
  switch_license --"CC BY or<br>CC BY-SA"--> same["<em>Same license for DB</em>"]

  sa --"Neither"--> cc0_db
  sa --"Attribution &<br>Indicate license &<br>Anti-DRM &<br>Copyleft"--> odbl["ODbL 1.0"]
